The role of geometry in the chain separation distribution of links in protein contact networks

### 2020-05-26 paper/main.tex

\red{With helices ranging in length from a minimum of about 2 to 3 turns or 7-11 residues up to beyond 30 residues \cite{manning1988circular}, with the (rarer) longer helices accounting relatively for more connections, sequence separations of 20 to 50 seem likely to result from connections between subsequent helices. (I guess we could test this somehow, though i'm not sure how I would go about interpreting the results. I think the PDB files contain information on which residues belong to which structure, so we could do an analysis of how many connections of $s>4$ there are a) within one helix b) between a helix and the following loop) c) between 2 subsequent helices etc. Where I would count helices and sheets the same...)}
\begin{figure}[h]
        \centering
	\includegraphics[width=0.35\textwidth]{paper/figures/neighbouring_helices.pdf}
        \caption{The average length of a link between two subsequent helices is the average length of a helix plus the average length of a loop. \red{Does this suffice or do we want a fancy 3d image of this? If so, do you have the software that makes those?}
        }
        \label{fig:bumpexplain}
\end{figure}

For a more direct comparison between simulation results and PRN's, we have generated a sample of intrinsically disordered proteins \red{cite something}, that also do not contain secondary structures. The results are shown in Fig.~\ref{fig:sdd_idp}.
\red{Maybe write how we found those samples and how many there are and all that.}
We find that the over representation of connections with $20<s<50$ disappears, further confirming our expectation.
\begin{figure}[h]
        \centering
	\includegraphics[width=0.45\textwidth]{figures/idp_200.pdf}
	\includegraphics[width=0.45\textwidth]{figures/idp_400.pdf}
        \caption{Sequence distance distributions for intrinsically disordered proteins and simulated folds for a) Chain lengths around 200 b) chain lengths around 400. The theoretical curve for 2D rings works okay for 3D open chains (because the rings are closed sequence separations of more than N/2 are a bit meaningless and thus not accounted for) The parameters used in this are $A=80$ and $\alpha=0.2$ in a) and $0.3$ in b). I thought in b) it makes sense for $\alpha$ to be larger as there are even larger exponents $k$ coming into relevance. I'm not sure we want to use that whole $\alpha$ argument too much though as it's a but ad-hoc and would be nicer to just fit $A$ and be done with it, but then it may look a bit more off.
        }
        \label{fig:sdd_idp}
\end{figure}
\red{Anything else we want to say to that? Maybe write down more fit parameters? Somehow i'm reluctant to do that because we wouldnt learn anything new from that, that i can think of and it doesnt look great number wise i guess.}
\section*{Conclusion}
We have demonstrated in analytical calculations and direct simulations how a random linking models with geometric constraints, such as the ones introduced in \cite{molkenthin2016scaling, molkenthin2020self} generate sequence distance distributions, that resemble a power-law, similar to the one found in folded proteins. This feature of protein network structure has been used to generate protein-like structures in \cite{bartoli2008effect}.

>>>>

\red{With helices ranging in length from a minimum of about 2 to 3 turns or 7-11 residues up to beyond 30 residues \cite{manning1988circular}, with the (rarer) longer helices accounting relatively for more connections, sequence separations of 20 to 50 seem likely to result from connections between subsequent helices. (I guess we could test this somehow, though i'm not sure how I would go about interpreting the results. I think the PDB files contain information on which residues belong to which structure, so we could do an analysis of how many connections of $s>4$ there are a) within one helix b) between a helix and the following loop) c) between 2 subsequent helices etc. Where I would count helices and sheets the same...)}
\begin{figure}[h]
        \centering
	\includegraphics[width=0.35\textwidth]{paper/figures/neighbouring_helices.pdf}
        \caption{The average length of a link between two subsequent helices is the average length of a helix plus the average length of a loop. \red{Does this suffice or do we want a fancy 3d image of this? If so, do you have the software that makes those?}
        \green{I can make a prettier thing provided we can actually make this argument based on the pdb data. I still struggle make the pdb data comply.}
        }
        \label{fig:bumpexplain}
\end{figure}

For a more direct comparison between simulation results and PRNs, we have generated a sample of intrinsically disordered proteins \red{cite something}, that also do not contain secondary structures. The results are shown in Fig.~\ref{fig:sdd_idp}.
In order to compare the 3D protein model simulations to PRNs, we selected protein structure lengths based on frequency and comparable length to the simulations. Fig.~\ref{fig:pdb_stats} shows that most proteins in the PDB have chain lengths between 100-400 protein residues. Therefore using example lengths of 200 and 400 residues seemed appropriate. From each of the over 20000 structures around the 200 and 400 residue length, we made the following choices for selecting different types of structures:
\red{now fill in what we actually did once we finalise this}.

\begin{figure}[h]
        \centering
	\includegraphics[width=0.45\textwidth]{figures/pdb_statistics.pdf}
        \caption{Statistics on all PDB entries based on the number of amino acid residues, with most entries found in the interval of 100-400 residues as of May 26th 2020. 
        }
        \label{fig:pdb_stats}
\end{figure}
\red{Maybe write how we found those samples and how many there are and all that.}
We find that the over representation of connections with $20<s<50$ disappears, further confirming our expectation.
\begin{figure}[h]
        \centering
	\includegraphics[width=0.45\textwidth]{figures/idp_200.pdf}
	\includegraphics[width=0.45\textwidth]{figures/idp_400.pdf}
        \caption{Sequence distance distributions for intrinsically disordered proteins and simulated folds for a) Chain lengths around 200 b) chain lengths around 400. The theoretical curve for 2D rings works okay for 3D open chains (because the rings are closed sequence separations of more than N/2 are a bit meaningless and thus not accounted for) The parameters used in this are $A=80$ and $\alpha=0.2$ in a) and $0.3$ in b). I thought in b) it makes sense for $\alpha$ to be larger as there are even larger exponents $k$ coming into relevance. I'm not sure we want to use that whole $\alpha$ argument too much though as it's a but ad-hoc and would be nicer to just fit $A$ and be done with it, but then it may look a bit more off.
        }
        \label{fig:sdd_idp}
\end{figure}
\red{Anything else we want to say to that? Maybe write down more fit parameters? Somehow i'm reluctant to do that because we wouldnt learn anything new from that, that i can think of and it doesnt look great number wise i guess.}
\section*{Conclusion}
We have demonstrated in analytical calculations and direct simulations how a random linking models with geometric constraints, such as the ones introduced in \cite{molkenthin2016scaling, molkenthin2020self} generate sequence distance distributions, that resemble a power-law, similar to the one found in folded proteins. This feature of protein network structure has been used to generate protein-like structures in \cite{bartoli2008effect}.
